---
title: "EDA"
author: "Agata Kaczmarek & Agata Makarewicz & Jacek Wiśniewski"
date: "27 03 2021"
output: 
  html_document:
  theme: united

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(knitr)
library(splitstackshape)
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(GGally)
library(DataExplorer)
library(gridExtra)
```

### Dataset

At the beginning, we would like to see how our data looks like eg. what are column names, values and the structure of whole data set. 

```{r data}
data <- read.csv('phones.csv')
knitr::kable(head(data))
```

##### Structure 

```{r str}
# structure
str(data)
```

##### Summary

Here there are are a basic statistics of numeric values, eg we can see that the biggest difference in values between min and max is for *battery_mAh*. Mean for *back_camera_mpix* is two times bogger than for *front_camera_mpix*.

```{r summary}
summaryDF <- function(indf) {
  require(splitstackshape)
  temp <- data.table(summary(indf))[, c("V2", "N"), with = FALSE]
  dcast.data.table(cSplit(temp, "N", ":")[!is.na(N_1)],
                   N_1 ~ V2, value.var = "N_2")
}


kable(summaryDF(data[,-c(1,2)])  %>% 
        arrange(front_camera_mpix) %>%
        rename(stats = N_1))
```

### Discrete variables 

Below we can see how many times each brand occurs in our set. Xiaomi and Apple are the most popular and there are a few which rarely occurs.

```{r brand}
data %>% 
	group_by(brand) %>% 
	summarise(count = n()) %>%
	ggplot(aes(x = reorder(brand,(count)), y = count)) + 
		geom_bar(stat = 'identity') +
    labs(x = 'brand') +
    coord_flip()
```

`name` variable:

```{r}
a <- c("Unique","Total","Percent")
b <- c(length(unique(data$name)),dim(data)[1], length(unique(data$name))/dim(data)[1])
t <- data.frame(t(b), row.names = NULL)
colnames(t) <- a
row.names(t) <- 'name'
kable(t)
```
### Numerical variables

For various variables shown below, there are a lot of different values for each, there are no clearly visible dominant values. On chart about screen diagonal, we can see pick moved to the right, the majority of ohones does have diaonal greater than 4.

```{r hist, fig.width=15}

n_bin <- floor(sqrt(dim(data)[1]))
bin_width <- floor((max(data$back_camera_mpix, na.rm=TRUE)-min(data$back_camera_mpix, na.rm=TRUE))/n_bin)
plot1 <- ggplot(data, aes(x=back_camera_mpix)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "Back Camera Megapixels", y = "Density")

bin_width <- floor((max(data$front_camera_mpix, na.rm=TRUE)-min(data$front_camera_mpix, na.rm=TRUE))/n_bin)
plot2 <- ggplot(data, aes(x=front_camera_mpix)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "Front Camera Megapixels", y = "Density")

bin_width <- floor((max(data$battery_mAh, na.rm=TRUE)-min(data$battery_mAh, na.rm=TRUE))/n_bin)
plot3 <- ggplot(data, aes(x=battery_mAh)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "Battery capacity in mAh", y = "Density")

bin_width <- round((max(data$ram_gb, na.rm=TRUE)-min(data$ram_gb, na.rm=TRUE))/n_bin, 1)
plot4 <- ggplot(data, aes(x=ram_gb)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "RAM memory in GB", y = "Density")

bin_width <- floor((max(data$flash_gb, na.rm=TRUE)-min(data$flash_gb, na.rm=TRUE))/n_bin)
plot5 <- ggplot(data, aes(x=flash_gb)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "Flash memory storage in GB", y = "Density")

bin_width <- round((max(data$diag, na.rm=TRUE)-min(data$diag, na.rm=TRUE))/n_bin, 1)
plot6 <- ggplot(data, aes(x=diag)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "Screen diagonal", y = "Density")

bin_width <- floor((max(data$height_px, na.rm=TRUE)-min(data$height_px, na.rm=TRUE))/n_bin)
plot7 <- ggplot(data, aes(x=height_px)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "Screen height in Megapixels", y = "Density")

bin_width <- floor((max(data$width_px, na.rm=TRUE)-min(data$width_px, na.rm=TRUE))/n_bin)
plot8 <- ggplot(data, aes(x=width_px)) +   
   geom_histogram(aes(y=..density..),colour="darkblue",fill="lightblue", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "steelblue2", size=1.5)+
   labs(x = "Screen width in Megapixels", y = "Density")

grid.arrange(plot1, plot2, ncol=2)
grid.arrange(plot3, plot4, ncol=2)
grid.arrange(plot5, plot6, ncol=2)
grid.arrange(plot7, plot8, ncol=2)
```

### Target analysis

As we see below, our target seems as if it is from normal distribution, except the fact that it is shifted to the left, the majority of phones cost less than 2500 zł.

```{r target, fig.width=15}
bin_width <- floor((max(data$price, na.rm=TRUE)-min(data$price, na.rm=TRUE))/n_bin)
plot1 <- ggplot(data, aes(x=price)) +   
   geom_histogram(aes(y=..density..),colour="darkorchid",fill="thistle2", binwidth = bin_width, boundary = 0) + 
   geom_density(aes(y=..density..), colour = "plum4", size=1.5)+
   labs(x = "Price in PLN", y = "Density") +
    ggtitle('Target variable distribution')


plot2 <- ggplot(data, aes(x="price", y=price)) +
  geom_boxplot(color="orange", fill="khaki1",lwd=1) + coord_flip()

grid.arrange(plot1, plot2, ncol=2)
```

#### Most popular brands vs target 

Below we see differences in prices of phons from four main brands - Apple, Huawei, Samsung and Xiaomi. Clearly visible is that Xiaomi are generally less expensive than the other brands.

```{r violin}
top_brands <- data %>% 
  group_by(brand) %>% 
	summarise(count = n()) %>%
  arrange(desc(count)) %>%
  select(brand) %>%
  head(4)

data %>% 
  filter(brand %in% top_brands$brand) %>%
  ggplot(aes(x=brand, y=price)) + 
  geom_violin(aes(fill=brand)) +
  geom_boxplot(width = 0.08) +
  scale_fill_brewer(palette="Accent") + theme_classic()+
  theme(legend.position = "none") +
    ggtitle('Target variable distribution for 4 most common brands')

```

### Correlations

#### Height & width

```{r}
data[is.na(data)]<-0
ggplot(data)+
  geom_count(mapping = aes(x=width_px, y=height_px))+
  labs(x= "Width in px", y="Height in px", size = 'number of phones')

```

Above we see that there is nearly linear dependence between *height_px* and *width_px*.

#### All numerical variables

```{r corr, fig.width=12}
data[is.na(data)]<-0
ggcorr(data, method = c("everything", "pearson"), limits = c(0,1), label = TRUE, layout.exp = 2, hjust = 0.75)
```

Correlation between *flash_gb* and *price* seems to be highest, and between *ram_gb*, *width_px*, *height_px* and target seems also to be high.

#### Relationships between variables

```{r, fig.height=11, fig.width=15}
data2<-data
data(data2, package = "reshape")
ggpairs(
  data2[, c(3, 4,5,8, 9, 10)],
  upper = list(continuous = "density", combo = "box_no_facet"),
  lower = list(continuous = "points", combo = "dot_no_facet")
)

```

### Summary
Dataset phones has few missing values, which will make modelling easier. Before creating first models, there are some issues to consider.

* Variables describing phone's size are strongly correlated. Maybe we should resign from two of them.

* There are some brands that are represented by few phones. Maybe we should merge them with label 'other'.